Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-12-29 grove-paper.tex
minor clean up, mostly graph decomp

%% Packages
\usepackage{booktabs}
\usepackage[rule=false]{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}
\usepackage{array}
\usepackage{todonotes}
\usepackage[ruled]{algorithm2e}

@@

\DeclareMathOperator{\arity}{\text{arity}}
\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
\DeclareMathOperator{\deleted}{\text{deleted}}
\DeclareMathOperator{\dom}{\text{dom}}
\DeclareMathOperator{\edges}{\text{edges}}
\DeclareMathOperator{\known}{\text{known}}
\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
\DeclareMathOperator{\multiparents}{\text{multiparents}}
\DeclareMathOperator{\orphans}{\text{orphans}}
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
\DeclareMathOperator{\reachable}{\text{reachable}}
\DeclareMathOperator{\seen}{\text{seen}}
\DeclareMathOperator{\unseen}{\text{unseen}}
\DeclareMathOperator{\vertexes}{\text{vertexes}}
\DeclareMathOperator{\vertex}{\text{vertex}}
\DeclareMathOperator{\publish}{\text{publish}}
\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
\DeclareMathOperator{\downIndex}{\text{downIndex}}
\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}

@@

We define the semantics of graph actions via the following transition relation
between graphs.
\[
  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
\]
In words, applying graph action $\left(s, \e\right)$ to graph $G$ results in the
updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
$G$.

@@

Translating a graph into expressions in the calculus allows us to model
user interaction as a stream of edits as described in \autoref{sub:User
Actions}. Our decomposition algorithm partitions the graph into disjount
connected components. We encode cycles by resolving all but one vertex of each
multi-edge into a by-id reference to the one vertex.

\begin{algorithm}[t]
  \caption{Graph Decomposition}
  \label{fig:Graph Decomposition}

  \DontPrintSemicolon
  \SetAlgoNoLine
  \SetKwFunction{reachable}{reachable}
  \SetKwProg{Fn}{function}{}{end}

  \KwIn{Graph $G$}
  \KwOut{A partitioning of the edges of $G$}

  $live := \{\e : G(\e) = \Plus\}$\;
  $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
  $D := \{\reachable(\e) : \e \notin live\}$\;
  $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
  $MP := \{\reachable(\e) : \e \in multiparent\}$\;
  $SC := \varnothing$\;
  $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
  \While{$remaining \ne \varnothing$}{
    $(root, cycle) := \reachable(\min(remaining))$\;
    $remaining := remaining \setminus \{root\} \setminus cycle$\;
    $SC := SC \cup \{(root, cycle)\}$\;
  }
  \Return{$(R, D, MP, SC)$}\;
  \;

  \Fn{$\reachable\ ((u, v, p, v'))$}{
    $children := \{(u', v', p', v'') \in live\}$\;
    $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
    \Return{$((u, v, p, v'), descendants)$}
  }
\end{algorithm}

The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
\wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
expressions that involve vertices with multiple parents. $D$ contains deleted
expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
$SC$ contains the deleted expressions that do not involve multiple parents but
do contain cycles.

A simple cycle is a set of vertices connected by a closed path, consisting
of:

1) two or more vertices, all descendants of each other, and
2) the vertices of any non-multiparent subtrees reachable from the core, and
not reachable from the root vertex.

To determine the simple cycles of a graph, we:

1) check if any vertex in the graph is a descendant of itself or some other
vertex that is (a descendant of itself),
2) find the least vertex in the core of this new(ly discovered) cycle, and
3) return a tree connecting every vertex reachable from the least (vertex).

The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.

<Mathematical definition of above.>

>>>>

%% Packages
\usepackage{booktabs}
\usepackage[rule=false]{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}
\usepackage{array}
\usepackage{todonotes}
% \usepackage[ruled]{algorithm2e}

@@

\DeclareMathOperator{\arity}{\text{arity}}
\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
\DeclareMathOperator{\dom}{\text{dom}}
\DeclareMathOperator{\edges}{\text{edges}}
\DeclareMathOperator{\known}{\text{known}}
\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
\DeclareMathOperator{\multiparents}{\text{multiparents}}
\DeclareMathOperator{\orphans}{\text{orphans}}
\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
\DeclareMathOperator{\reachable}{\text{reachable}}
\DeclareMathOperator{\seen}{\text{seen}}
\DeclareMathOperator{\unseen}{\text{unseen}}
\DeclareMathOperator{\vertexes}{\text{vertexes}}
\DeclareMathOperator{\vertex}{\text{vertex}}
\DeclareMathOperator{\publish}{\text{publish}}
\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
\DeclareMathOperator{\downIndex}{\text{downIndex}}
\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}

% graph decomposition operators
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\multiParented}{\text{multiParented}}
\DeclareMathOperator{\singleParented}{\text{singleParented}}
\DeclareMathOperator{\deleted}{\text{deleted}}

@@

We define the semantics of graph actions via the following transition relation
between graphs.
\[
  G \overset{(s,\e)}{\longrightarrow} G\left[ \e \mapsto s \sqcup G(\e) \right\}
\]
Applying graph action $\left(s, \e\right)$ to graph $G$ results in the
updated graph $G' = G\left[\e \mapsto s \sqcup G(\e) \right]$, wherein the edge
state associated with edge $\e$ in $G$ becomes the join of $s$ with the state of
$\e$ in $G$, and $G'(\e') = G(\e')$ for all other $\e' \ne \e$ in the domain of
$G$.

@@

% To render the graph as disjoint fragments of program code, we decompose the
% graph into an equivalent representation using trees. The result is a 4-tuple,
% where the first component is a tree, the second and third components are sets of
% trees, and the fourth component is a set of cycles.

% 1. ($O(n \log n)$ for maps, $O(n)$ for hash table) linear scan to partition vertices according to # of parents: multiple parents, single parents, no parents.

% 2. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from root (and remove from single-parented set)

% 3. ($n \log n$?) scan to discover edges leading to single-parented vertices reachable from multi-parented vertices (and remove from single-parented set).

% 4. ($n \log n$?) scan to discover edges leading to single-parented [deleted]

% 5. simple cycles—reverse traversal, then forward traversal. Here's where we want to prove some things (e.g., dangling vertices are not deleted)

% tree ::= Vertex (Constructor) | Ref (Vertex.Id) | Conflict (tree list)

%
% ... () of sets of
% (poly)trees) wherein each multi-edge has been resolved into a simple edge with
% one or more symbolic references to it.

% Our decomposition algorithm first separates the graph into three disjoint sets of
% vertices according to the number of parents and presence of simple cycles.
% %
% The first set is the single-parented vertices that have the root as an ancestor.
% The second set is the vertices that have a multi-parented vertex as an ancestor.
% The third

% We then derive equivalent tree representations of connected components by tracing
% the ancestry, taken as the transitive closure of the childhood relation, of the
% vertices in each set.

% set contains the multi-parented vertices and
% their non-multiparented descendants. The second set contains the vertices with
% no parents and their non-multiparented descendants. We can determine both of
% these sets with a single pass over all of the live vertices in the graph. The
% third set contains the single-parented vertices that are descendants of the root
% vertex.

% The first set contains all of the
% vertices that have the root as an ancestor.

% : vertices with multiple parents, vertices with no parents,
% single-parented vertices reachable from the root, and everything else.

% separate it into 
% components that only touch  detect any vertices that either have multiple parents or no parents
% at all.

% partition its vertices into three disjoint sets: multi-parented, deleted,
% and any remaining vertices participating in a live edge. We then partition the
% graph into four non-overlapping sets of trees: multi-parented, single-parented
% reachable from root, deleted, and simple cycles.

%  detect any multi-parented
% vertices and use them to partition the remaining vertices of the graph into .

% Our decomposition algorithm partitions the graph into trees of
% connected components. Cycles are resolved by replacing all but one vertex of
% each multi-edge with a by-id reference.

% Translating a graph into expressions in the calculus allows us to model user
% interaction as a stream of edits as described in \autoref{sub:User Actions}.

% \begin{algorithm}[t]
%   \caption{Graph Decomposition}
%   \label{fig:Graph Decomposition}

%   \DontPrintSemicolon
%   \SetAlgoNoLine
%   \SetKwFunction{reachable}{reachable}
%   \SetKwProg{Fn}{function}{}{end}

%   \KwIn{Graph $G$}
%   \KwOut{A partitioning of the edges of $G$}

%   $live := \{\e : G(\e) = \Plus\}$\;
%   $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
%   $D := \{\reachable(\e) : \e \notin live\}$\;
%   $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
%   $MP := \{\reachable(\e) : \e \in multiparent\}$\;
%   $SC := \varnothing$\;
%   $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
%   \While{$remaining \ne \varnothing$}{
%     $(root, cycle) := \reachable(\min(remaining))$\;
%     $remaining := remaining \setminus \{root\} \setminus cycle$\;
%     $SC := SC \cup \{(root, cycle)\}$\;
%   }
%   \Return{$(R, D, MP, SC)$}\;
%   \;

%   \Fn{$\reachable\ ((u, v, p, v'))$}{
%     $children := \{(u', v', p', v'') \in live\}$\;
%     $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
%     \Return{$((u, v, p, v'), descendants)$}
%   }
% \end{algorithm}

% The result of decomposition is a 4-tuple $(e, MP, D, SC) \in Exp \times
% \wp({Exp}) \times \wp({Exp}) \times \wp({Exp})$. Expression $e$ covers the
% portion of the graph reachable from the root vertex $v_0$. $MP$ contains the
% expressions that involve vertices with multiple parents. $D$ contains deleted
% expressions, i.e., those not reachable from $v_0$, that do not contain cycles.
% $SC$ contains the deleted expressions that do not involve multiple parents but
% do contain cycles.

% A simple cycle is a set of vertices connected by a closed path, consisting
% of:

% 1) two or more vertices, all descendants of each other, and
% 2) the vertices of any non-multiparent subtrees reachable from the core, and
% not reachable from the root vertex.

% To determine the simple cycles of a graph, we:

% 1) check if any vertex in the graph is a descendant of itself or some other
% vertex that is (a descendant of itself),
% 2) find the least vertex in the core of this new(ly discovered) cycle, and
% 3) return a tree connecting every vertex reachable from the least (vertex).

% The least vertex of a cycle is the one that was created first. To ensure consistent behavior, the search for reachable vertices begins at the least vertex in a cycle.

% <Mathematical definition of above.>
